Substitution rate response to changes in effective population size and protein expression level.

### 2020-06-16 manuscript/supp-mat.tex
Manuscript formatted to MBE guidelines

\documentclass{article}
\usepackage{url}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{caption}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{xfrac, nicefrac}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage[margin=60pt]{geometry}
\pdfinclusioncopyfonts=1
\captionsetup{width=0.85\textwidth}
\include{notations}

@@

\section{$\bm{\omega}$ elasticity after a change in $\bm{\Ne}$}

@@

p_{\text{fix}}(x, \dx) &= \frac{ 2 s(x, \dx)}{1 - \e^{-4\Ne s(x, \dx)}}, \\

@@

\subsection{$\bm{\omega}$ elasticity after a change in $\bm{\Ne}$}

@@

Together, these approximations leads to the following elasticity in equilibrium $\omega$ after change in $\Ne$ as:

@@

A first model is to assume that the fitness is equal to the folded fraction \cite{Goldstein2013}:

@@

Finally, $\omega$ elasticity after a change in $\Ne$ is simply:

@@

A slight variation is to assume that the selective cost itself is proportional to the total amount of misfolded protein. For a given protein with expression level $y$:

@@

\frac{\partial \ln (f(x))}{\partial x} &\simeq& -A \beta \gamma \Nsite y  e^{\beta(\alpha + \gamma \Nsite x)}

@@

@@

4 \Ne \frac{\partial \ln (f(x))}{\partial x} &\propto& -4 \Ne y e^{\beta(\alpha + \gamma \Nsite x)}

@@

Thus, under this model, an increase in the effective population size $\Ne$ or the expression level $y$ both have the same effect on the molecular evolutionary process undergone by the protein.
Moreover, $\omega$ elasticity after a change in $\Ne$ is the same as before:

@@

@@

Moreover, $\omega$ elasticity after a change in $\Ne$ is again the same as before:

@@

This models assumes that the :

@@

 \item the log-fitness is proportional to the ratio of the \emph{total} cost of expression over the benefit contributed by the protein (Beaulieu et al).

@@

Alternative cost-exp models could also be used, allowing for a non-linear cost function for expression or for some elasticity of the realized equilibrium expression level, as a function of the number of mutations (Gout et al). Under these models, the strength of selection is still expected to be an increasing function of $y$, although not linear, e.g.

@@

Finally, applying the main result of this work to the present case allows us to express the elasticity of $\omega$ as a function of $\Ne$ as:

@@

Based on empirical estimates obtained from Zhang et al. The mean fraction of hydrophobic residues at the surface of proteins is $0.22 \pm 0.06$. With $n=100$ residues, this makes $22 \pm 6$. The mean value for $E_{ij}$ is 7 kT, with a standard deviation of $\sigma = 1.8$ kT. Assuming that this standard deviation of $\pm 1.8$ kT is contributed by $\pm 6$ mutations gives $\gamma = 1.8 / 6 = 0.3$ kT or $0.18$ kcal per mole. Also, with $x=0.22$, $\lambda \simeq 4$, and thus $\chi = 6  / 30 = 0.2$, thus a much stronger response than under the model based on conformational stability.

\section{Simulation of protein-coding DNA sequences evolution}
\subsection{Probability of folding using the 3d structure of protein}
We simulated substitutions in the protein phosphatase ($\Nsite=300$ codon sites).
From a DNA sequence $\ci$ after $t$ substitutions, we compute the free energy of the folded state $G_{\mathrm{F}}\left(\ci\right)$, using the $3$-dimensional structure of the folded state and pair-wise contact energies between neighboring amino-acid residues:
\begin{equation}
G_{\mathrm{F}}\left(\ci\right) = \sum_{1 \leq \site \leq \Nsite} \sum_{r \in \mathcal{N}(\site)} I \left(\ci(\site), \ci(r) \right),
\end{equation}
where $I(a,b)$ is the pair-wise contact energies between amino-acid $a$ and $b$, using contact potentials estimated by Miya-zawa and Jernigan, and $\mathcal{N}(\site)$ are the neighbor residues of site $\site$ (closer than $7\angstrom$) in the $3$D structure.\\
The free energy of unfolded states $G_{\mathrm{U}}\left(\ci\right)$ is approximated using $55$ decoy $3$D structures that supposedly represent a sample of possible unfolded states:
\begin{equation}
G_{\mathrm{U}}\left(\ci\right) = \langle G\left(\ci\right) \rangle - kT \ln (1.0\mathrm{E}^{160}) - \frac{2 \left[ \langle G\left(\ci\right)^2 \rangle - \langle G\left(\ci\right) \rangle^2\right] }{kT}
\end{equation}
where the average $\langle . \rangle$ runs other the $55$ decoy $3$D structures, and $k$ is the Boltzmann constant and $T$ the temperature in Kelvin.\\
From the energy of folded and unfolded states, we can compute the difference in free energy between the states:
\begin{equation}
\deltaG\left(\ci\right) = G_{\mathrm{F}}\left(\ci\right) - G_{\mathrm{U}}\left(\ci\right)
\end{equation}
\begin{center}
 \includegraphics[width=165mm] {artworks/ModelSimuFold.pdf}
\end{center}

\subsection{Simulated $\bm{\omega}$ elasticity to changes in $\bm{\Ne}$}
 \includegraphics[width=0.8\textwidth] {artworks/SimuDfe-Elasticity.pdf}
\textbf{$\bm{\omega}$ elasticity with gamma distributed selection coefficient}.
$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
In the model of gamma distributed fitness effect, $\omega$ at equilibrium is strongly dependent on log-$\Ne$ where the slope correlation is proportional to the inverse of the shape parameter of the gamma distribution \cite{Welch2008}.
 \includegraphics[width=0.8\textwidth] {artworks/SimuProfile-Elasticity.pdf}
\textbf{$\bm{\omega}$ elasticity with amino-acid fitness profiles}.
$\omega$ at equilibrium as a function of $\bm{\Ne}$ (relative).
For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
In the model of site-wise amino-acid fitness profiles taken from \cite{Bloom2017}, $\omega$ at equilibrium is strongly dependent on log-$\Ne$. 
 \includegraphics[width=0.8\textwidth] {artworks/SimuStab-Elasticity.pdf}
\textbf{$\bm{\omega}$ elasticity with additive free energy of folding}.
$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
The fixed parameters are $\alpha=-118$, $\gamma=1$, $\Nsite=300$, $\beta=1.686$.
The simulations of our additive free energy model match the theoretical prediction that the slope of the linear relation (dashed line) is equal to $\beta \Nsite \gamma)^{-1} = 0.00198 \simeq = 0.00199$.
%The non-monotony is suspected to be due to the discrete number of sites and states, such that the changes in $\deltaG$ after a mutation is either $-1$, $0$ or $1$. Such non-monotony is not observed with the Grantham model, in which the $\omega$ is lower and the slope of elasticity lower, closer to the empirical 3D model of Golstein \& Pollock. 
%\begin{center}
% \includegraphics[width=0.8\textwidth] {artworks/SimuStab-Grantham-Beta-Elasticity.pdf}
%\end{center}
%\textbf{Effect of $\bm{\beta}$ on the elasticity}.
%$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
%For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
%The fixed parameters are $\alpha=-118$, $\gamma=1$, $\Nsite=300$, and for each non-optimal amino-acid, $\gamma$ is scaled by the Grantham distance to the optimal amino-acid. $\beta=1.686$ are given in the legend.
%Increasing $\beta$ increases the slope of the $\omega$-$\Ne$ relationship, as predicted in our theoretical model.
%\begin{center}
% \includegraphics[width=0.8\textwidth] {artworks/SimuStab-GammaStd-Elasticity.pdf}
%\end{center}
%\textbf{$\bm{\omega}$ elasticity with additive free energy of folding}.
%$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
%For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
%The parameters are $\alpha=-118$, $\Nsite=300$, $\beta=1.686$ and each site has it's own gamma distributed $\gamma$ with mean $1$ and standard deviation given in the legend.
%Increasing the variance of $\gamma$ increases $\omega$, by shifting the equilibrium to higher $x\eq$ since more unstable sites with low $\gamma$ are fixed before reaching sensible deleterious selection coefficient against unstable mutations. Once many sites are unstable, the $\omega$ is higher since non-synonymous mutations between unstable states are effectively neutral. However the slope of the $\omega$-$\Ne$ relationship is not sensibly changed.
\subsection{Simulated relaxation time of $\bm{\omega}$}
 \includegraphics[width=0.8\textwidth] {artworks/Relaxation-Stability.pdf}
\textbf{Relaxation time of $\bm{\omega}$ dependence on $\Nsites$}.
$\omega$ Relaxation after a brutal change in $\Ne$, the left and right panel correspond to low $\Ne$ ($1e^{5}$) and the middle panel corresponds to high $\Ne$ ($2e^{6}$). 
Solid line corresponds to the average over replicates ($r$) and the shaded area correspond to the $90\%$ interval among replicates. 
The mutation rate ($\mu$) is $1e{-8}$ per year per site, and the total time of the computation is $900$ million years.
$\beta=1.686$, $\gamma=0.2$ and $\alpha=-10$ for all simulations. The number of sites is changed from $\Nsite=15$ to $\Nsite=158$, and the number of replicates is changed accordingly such that the total number of sites ($\Nsite*r$) is kept constant.
Increasing $\Nsite$ implies a higher $\omega$ at equilibrium, a lower elasticity of the $\omega$ to changes in $\Ne$ and a higher rate of relaxation.
 \includegraphics[width=0.8\textwidth] {artworks/Relaxation-Stability-Gamma.pdf}
\textbf{Relaxation time of $\bm{\omega}$ dependence on $\Nsites$, while correction for $\gamma$}.
$\omega$ Relaxation after a brutal change in $\Ne$, the left and right panel correspond to low $\Ne$ ($1e^{5}$) and the middle panel corresponds to high $\Ne$ ($2e^{6}$). 
Solid line corresponds to the average over replicates ($r$) and the shaded area correspond to the $90\%$ interval among replicates. 
The mutation rate ($\mu$) is $1e{-8}$ per year per site, and the total time of the computation is $900$ million years.
$\beta=1.686$, $\alpha=-10$ for all simulations. The number of sites is changed from $\Nsite=15$ to $\Nsite=158$, and the number of replicates is changed accordingly such that the total number of sites ($\Nsite*r$) is kept constant.
Moreover, $\gamma$ is changed according to $\Nsite$ such that the product $\gamma\Nsite$ is kept constant, thus the  elasticity of the $\omega$ to changes in $\Ne$ is kept constant.
Increasing $\Nsite$ implies a higher $\omega$ at equilibrium, and a higher rate of relaxation.
 \includegraphics[width=0.8\textwidth] {artworks/Relaxation-Stability-Grantham-Alpha-Gamma.pdf}
\textbf{Relaxation time of $\bm{\omega}$ for the Grantham model}.
$\omega$ Relaxation after a brutal change in $\Ne$, the left and right panel correspond to low $\Ne$ ($1e^{5}$) and the middle panel corresponds to high $\Ne$ ($2e^{6}$). 
Solid line corresponds to the average over replicates ($r$) and the shaded area correspond to the $90\%$ interval among replicates. 
The mutation rate ($\mu$) is $1e{-8}$ per year per site, and the total time of the computation is $900$ million years.
$\beta=1.686$, $\gamma=-10$ for all simulations. The number of sites is changed from $\Nsite=15$ to $\Nsite=158$, and the number of replicates is changed accordingly such that the total number of sites ($\Nsite*r$) is kept constant.
Moreover, $\gamma$ is changed according to $\Nsite$ such that the product $\gamma\Nsite$ is kept constant, thus the  elasticity of the $\omega$ to changes in $\Ne$ is kept constant.
Finally, $\alpha$ is changed according to $\Nsite$ and $\gamma$ such that the equilibrium value $x\eq$ is kept constant, by solving numerically equation \ref{eq:equilibrium}.
Increasing $\Nsite$ implies a higher rate of relaxation.

\subsection*{Distribution of fitness effects}
DNA mutations changing a genotype can result in a change of phenotype, and ultimately a change in fitness.
From a specific genotype, all the possible mutations thus result in a distribution a phenotypic effect (DPE) and fitness effects (DFE).
The DPE and DFE are not known a priori, but are the resulting consequence of the mutation-selection-drift balance.
Empirically, these distributions are of particular importance since they can be obtained experimentally or inferred with other data.
As an example, DFE can be inferred from polymorphism dataset \cite{Eyre-walker2007, Galtier2016}. 
Moreover, the distribution of $\deltadeltaG=\gamma \Nsite \dx$ of novel mutations can be obtained experimentally.

@@

\bibliographystyle{apalike}
\bibliography{refs-codons,refs-cds}
\end{document}

>>>>

\documentclass{article}
\usepackage[bookmarks]{hyperref}
\hypersetup{
    colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black,
	linktoc=all,
}

\usepackage[margin=60pt]{geometry}
\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{bm,bbold}
\pdfstringdefDisableCommands{%
	\renewcommand*{\bm}[1]{#1}%
	% any other necessary redefinitions 
}
\usepackage{xfrac, nicefrac}
\usepackage{lmodern}
\usepackage{natbib}
\pdfinclusioncopyfonts=1
\include{notations}

@@

\section{$\bm{\omega}$ susceptibility after a change in $\bm{\Ne}$}

@@

p_{\text{fix}}(x, \dx) &= \frac{ 1 - \e^{-2 s(x, \dx)}}{1 - \e^{-4\Ne s(x, \dx)}}, \\
 &\simeq \frac{ 2 s(x, \dx)}{1 - \e^{-4\Ne s(x, \dx)}}, \\

@@

\subsection{$\bm{\omega}$ susceptibility after a change in $\bm{\Ne}$}

@@

Together, these approximations leads to the following susceptibility in equilibrium $\omega$ after change in $\Ne$ as:

@@

A first model is to assume that the fitness is equal to the folded fraction \citep{Goldstein2013}:

@@

Finally, $\omega$ susceptibility after a change in $\Ne$ is simply:

@@

A slight variation is to assume that the selective cost itself is proportional to the total amount of misfolded protein  \citep{Drummond2005a, Wilke2006, Drummond2008, Serohijos2012}. For a given protein with expression level $y$:

@@

\frac{\partial \ln (f(x))}{\partial x} &\simeq& -A y \beta \gamma \Nsite e^{\beta(\alpha + \gamma \Nsite x)}.

@@

Under this model, the phenotype at equilibrium is given by:

@@

\lambda_{K}(x\eq) = 4\Ne y A \beta \gamma \Nsite e^{\beta(\alpha + \gamma \Nsite x\eq)} \text{ from eq. \ref{eq:equilibrium_lambda} and \ref{eq:df_dx}}.

@@

And the susceptibility of $\omega$ after a change in $\Ne$ is the same as before:

@@

Since $\Ne$ and $y$ are confounded factors, meaning they only appear in the equation as a product between the two, implicit derivation leads to the same result whenever the derivation is w.r.t $\Ne$ or $y$, leading to same compact susceptibility:
\frac{ \der \omega\eq}{\der \ln (y)} = \frac{ \der \omega\eq}{\der \ln (\Ne)} \simeq - \frac{1}{\beta \Nsite \gamma}.

@@

Moreover, $\omega$ susceptibility after a change in $\Ne$ is again the same as before:

@@

The cost-benefit argument \citep{Beaulieu2018} is based on two assumptions 

@@

 \item the log-fitness is proportional to the ratio of the \emph{total} cost of expression over the benefit contributed by the protein .

@@

Alternative cost-expression models could also be used, allowing for a non-linear cost function for expression or for some susceptibility of the realized equilibrium expression level, as a function of the number of mutations. Under these models, the strength of selection is still expected to be an increasing function of $y$, although not linear, e.g.

@@

Finally, applying the main result of this work to the present case allows us to express the susceptibility of $\omega$ as a function of $\Ne$ as:

@@

Based on empirical estimates \citep{Zhang2008}. The mean fraction of hydrophobic residues at the surface of proteins is $0.22 \pm 0.06$. With $n=100$ residues, this makes $22 \pm 6$. The mean value for $E_{ij}$ is 7 kT, with a standard deviation of $\sigma = 1.8$ kT. Assuming that this standard deviation of $\pm 1.8$ kT is contributed by $\pm 6$ mutations gives $\gamma = 1.8 / 6 = 0.3$ kT or $0.18$ kcal per mole. Also, with $x=0.22$, $\lambda \simeq 4$, and thus $\chi = 6  / 30 = 0.2$, thus a much stronger response than under the model based on conformational stability.

\section{Empirical estimation of $\bm{\chi}$}
Substitution rate as a function of expression level compiled by \citep{Zhang2015}.
\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|c|c|}
		\hline
		Type & Specie & $\hat{\chi}$ & $r^2$ \\ 
		\hline
		Plant & Oryza sativa & -0.008 & 0.047 \\ 
		Plant & Arabidopsis thaliana & -0.012 & 0.128 \\ 
		Archaea & Sulfolobus solfataricus & -0.037 & 0.097 \\ 
		Archaea & Thermococcus kodakarensis & -0.026 & 0.058 \\ 
		Fungi & Saccharomyces cerevisiae & -0.029 & 0.211 \\ 
		Fungi & Aspergillus nidulans & -0.034 & 0.124 \\ 
		Bacteria & Escherichia coli & -0.021 & 0.151 \\ 
		Bacteria & Bacillus subtilis & -0.046 & 0.151 \\ 
		Animal & Caenorhabditis elegans & -0.026 & 0.039 \\ 
		Animal & Drosophila melanogaster & -0.005 & 0.021 \\ 
		Animal & Mus musculus & -0.008 & 0.085 \\ 
		Animal & Homo sapiens & -0.004 & 0.031 \\ 
		\hline
	\end{tabular}
\end{table}
In \citep{Brevet2019}, the covariance matrix with $\ln(\Ne)$ and $\ln(\omega)$ as entries allow us to approximate $\chi$:
\begin{align}
	\widehat{\frac{\der \ln (\omega)}{\der \ln (\Ne)}} & = \dfrac{\text{Cov}[\ln(\omega), \ln(\Ne)]}{\text{Var}[\ln(\Ne)]}\text{ from linear regression.} \\
	\Rightarrow \widehat{\frac{\der \omega}{\omega \der \ln (\Ne)}} & = \dfrac{\text{Cov}[\ln(\omega), \ln(\Ne)]}{\text{Var}[\ln(\Ne)]} \\
	\Rightarrow \widehat{\chi} & \simeq \widehat{\omega} \dfrac{\text{Cov}[\ln(\omega), \ln(\Ne)]}{\text{Var}[\ln(\Ne)]} \\
	\Rightarrow \chi & \simeq 0.2 \dfrac{-0.45}{4.45} \\
	\Rightarrow \chi & \simeq - 0.02 
\end{align}
\section{Simulation using the 3d structure of protein}
We simulated substitutions in the protein phosphatase ($\Nsite=300$ codon sites).
From a DNA sequence $\ci$ after $t$ substitutions, we compute the free energy of the folded state $G_{\mathrm{F}}\left(\ci\right)$, using the $3$-dimensional structure of the folded state and pair-wise contact energies between neighboring amino-acid residues:
\begin{equation}
G_{\mathrm{F}}\left(\ci\right) = \sum_{1 \leq \site \leq \Nsite} \sum_{r \in \mathcal{N}(\site)} I \left(\ci(\site), \ci(r) \right),
\end{equation}
where $I(a,b)$ is the pair-wise contact energies between amino-acid $a$ and $b$, using contact potentials estimated by Miya-zawa and Jernigan, and $\mathcal{N}(\site)$ are the neighbor residues of site $\site$ (closer than $7\angstrom$) in the $3$D structure.\\
The free energy of unfolded states $G_{\mathrm{U}}\left(\ci\right)$ is approximated using $55$ decoy $3$D structures that supposedly represent a sample of possible unfolded states:
\begin{equation}
G_{\mathrm{U}}\left(\ci\right) = \langle G\left(\ci\right) \rangle - kT \ln (1.0\mathrm{E}^{160}) - \frac{2 \left[ \langle G\left(\ci\right)^2 \rangle - \langle G\left(\ci\right) \rangle^2\right] }{kT}
\end{equation}
where the average $\langle . \rangle$ runs other the $55$ decoy $3$D structures, and $k$ is the Boltzmann constant and $T$ the temperature in Kelvin.\\
From the energy of folded and unfolded states, we can compute the difference in free energy between the states:
\begin{equation}
\deltaG\left(\ci\right) = G_{\mathrm{F}}\left(\ci\right) - G_{\mathrm{U}}\left(\ci\right)
\end{equation}
\begin{center}
 \includegraphics[width=165mm] {artworks/ModelSimuFold.pdf}
\end{center}

\section{Simulated $\bm{\omega}$ susceptibility to changes in $\bm{\Ne}$}
\subsection{$\bm{\omega}$ susceptibility with gamma distributed selection coefficient}
\begin{center}
 \includegraphics[width=0.8\textwidth] {artworks/SimuDfe-Elasticity.pdf}
\end{center}
$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
In the model of gamma distributed fitness effect, $\omega$ at equilibrium is strongly dependent on log-$\Ne$ where the slope correlation is proportional to the inverse of the shape parameter of the gamma distribution \citep{Welch2008}.

\subsection{$\bm{\omega}$ susceptibility with amino-acid fitness profiles}.
\begin{center}
 \includegraphics[width=0.8\textwidth] {artworks/SimuProfile-Elasticity.pdf}
\end{center}
$\omega$ at equilibrium as a function of $\bm{\Ne}$ (relative).
For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
In the model of site-wise amino-acid fitness profiles taken from \citep{Bloom2017}, $\omega$ at equilibrium is strongly dependent on log-$\Ne$.

\subsection{$\bm{\omega}$ susceptibility with additive free energy of folding}.
\begin{center}
 \includegraphics[width=0.8\textwidth] {artworks/SimuStab-Elasticity.pdf}
\end{center}
$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
The fixed parameters are $\alpha=-118$, $\gamma=1$, $\Nsite=300$, $\beta=1.686$.
The simulations of our additive free energy model match the theoretical prediction that the slope of the linear relation (dashed line) is equal to $\beta \Nsite \gamma)^{-1} = 0.00198 \simeq = 0.00199$.
The non-monotony is suspected to be due to the discrete number of sites and states, such that the changes in $\deltaG$ after a mutation is either $-1$, $0$ or $1$. Such non-monotony is not observed with the Grantham model, in which the $\omega$ is lower and the slope of susceptibility lower, closer to the empirical 3D model.

%\subsection{Effect of $\bm{\beta}$ on the susceptibility}.
%\begin{center}
% \includegraphics[width=0.8\textwidth] {artworks/SimuStab-Grantham-Beta-Elasticity.pdf}
%\end{center}
%$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
%For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
%The fixed parameters are $\alpha=-118$, $\gamma=1$, $\Nsite=300$, and for each non-optimal amino-acid, $\gamma$ is scaled by the Grantham distance to the optimal amino-acid. $\beta=1.686$ are given in the legend.
%Increasing $\beta$ increases the slope of the $\omega$-$\Ne$ relationship, as predicted in our theoretical model.

%\subsection{Effect of site variance on the susceptibility}.
%\begin{center}
% \includegraphics[width=0.8\textwidth] {artworks/SimuStab-GammaStd-Elasticity.pdf}
%\end{center}
%$\omega$ at equilibrium as a function of $\bm{\Ne}$ (log scale).
%For each population size, $200$ simulations were performed and the average (solid line) and $90\%$ confidence interval (shaded area) are shown.
%The parameters are $\alpha=-118$, $\Nsite=300$, $\beta=1.686$ and each site has it's own gamma distributed $\gamma$ with mean $1$ and standard deviation given in the legend.
%Increasing the variance of $\gamma$ increases $\omega$, by shifting the equilibrium to higher $x\eq$ since more unstable sites with low $\gamma$ are fixed before reaching sensible deleterious selection coefficient against unstable mutations. Once many sites are unstable, the $\omega$ is higher since non-synonymous mutations between unstable states are effectively neutral. However the slope of the $\omega$-$\Ne$ relationship is not sensibly changed.

\section{Simulated relaxation time of $\bm{\omega}$}

\subsection{Relaxation time of $\bm{\omega}$ dependence on $\Nsites$}.
\begin{center}
 \includegraphics[width=0.8\textwidth] {artworks/Relaxation-Stability.pdf}
\end{center}

$\omega$ Relaxation after a brutal change in $\Ne$, the left and right panel correspond to low $\Ne$ ($1e^{5}$) and the middle panel corresponds to high $\Ne$ ($2e^{6}$). 
Solid line corresponds to the average over replicates ($r$) and the shaded area correspond to the $90\%$ interval among replicates. 
The mutation rate ($\mu$) is $1e{-8}$ per year per site, and the total time of the computation is $900$ million years.
$\beta=1.686$, $\gamma=0.2$ and $\alpha=-10$ for all simulations. The number of sites is changed from $\Nsite=15$ to $\Nsite=158$, and the number of replicates is changed accordingly such that the total number of sites ($\Nsite*r$) is kept constant.
Increasing $\Nsite$ implies a higher $\omega$ at equilibrium, a lower susceptibility of the $\omega$ to changes in $\Ne$ and a higher rate of relaxation.

\subsection{Relaxation time of $\bm{\omega}$ dependence on $\Nsites$, while correction for $\gamma$}.
\begin{center}
 \includegraphics[width=0.8\textwidth] {artworks/Relaxation-Stability-Gamma.pdf}
\end{center}
$\omega$ Relaxation after a brutal change in $\Ne$, the left and right panel correspond to low $\Ne$ ($1e^{5}$) and the middle panel corresponds to high $\Ne$ ($2e^{6}$). 
Solid line corresponds to the average over replicates ($r$) and the shaded area correspond to the $90\%$ interval among replicates. 
The mutation rate ($\mu$) is $1e{-8}$ per year per site, and the total time of the computation is $900$ million years.
$\beta=1.686$, $\alpha=-10$ for all simulations. The number of sites is changed from $\Nsite=15$ to $\Nsite=158$, and the number of replicates is changed accordingly such that the total number of sites ($\Nsite*r$) is kept constant.
Moreover, $\gamma$ is changed according to $\Nsite$ such that the product $\gamma\Nsite$ is kept constant, thus the  susceptibility of the $\omega$ to changes in $\Ne$ is kept constant.
Increasing $\Nsite$ implies a higher $\omega$ at equilibrium, and a higher rate of relaxation.

\subsection{Relaxation time of $\bm{\omega}$ for the Grantham model}.
\begin{center}
 \includegraphics[width=0.8\textwidth] {artworks/Relaxation-Stability-Grantham-Alpha-Gamma.pdf}
\end{center}
$\omega$ Relaxation after a brutal change in $\Ne$, the left and right panel correspond to low $\Ne$ ($1e^{5}$) and the middle panel corresponds to high $\Ne$ ($2e^{6}$). 
Solid line corresponds to the average over replicates ($r$) and the shaded area correspond to the $90\%$ interval among replicates. 
The mutation rate ($\mu$) is $1e{-8}$ per year per site, and the total time of the computation is $900$ million years.
$\beta=1.686$, $\gamma=-10$ for all simulations. The number of sites is changed from $\Nsite=15$ to $\Nsite=158$, and the number of replicates is changed accordingly such that the total number of sites ($\Nsite*r$) is kept constant.
Moreover, $\gamma$ is changed according to $\Nsite$ such that the product $\gamma\Nsite$ is kept constant, thus the  susceptibility of the $\omega$ to changes in $\Ne$ is kept constant.
Finally, $\alpha$ is changed according to $\Nsite$ and $\gamma$ such that the equilibrium value $x\eq$ is kept constant, by solving numerically equation \ref{eq:equilibrium}.
Increasing $\Nsite$ implies a higher rate of relaxation.

\section{Distribution of fitness effects}
DNA mutations changing a genotype can result in a change of phenotype, and ultimately a change in fitness.
From a specific genotype, all the possible mutations thus result in a distribution a phenotypic effect (DPE) and fitness effects (DFE).
The DPE and DFE are not known a priori, but are the resulting consequence of the mutation-selection-drift balance.
Empirically, these distributions are of particular importance since they can be obtained experimentally or inferred with other data.
As an example, DFE can be inferred from polymorphism dataset \citep{Eyre-walker2007, Galtier2016}. 
Moreover, the distribution of $\deltadeltaG$ for novel mutations can be obtained experimentally.

@@

\bibliographystyle{natbib}
\bibliography{references}
\end{document}
